---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
First, clean the working environment:
```{r echo=TRUE}
rm(list=ls(all=TRUE))
```
Next, download the data:
```{r echo=TRUE}
datafile <- "activity.zip"
if(!file.exists(datafile)){
	url <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
	download.file(
		url,
		destfile = datafile,
		method = "wget",
		quiet = TRUE
	)
	unzip(
		zipfile = datafile,
	)
}
```
Data description, from course website, as follows:
> This assignment makes use of data from a personal activity monitoring device. This device collects data at 5 minute intervals through out the day. The data consists of two months of data from an anonymous individual collected during the months of October and November, 2012 and include the number of steps taken in 5 minute intervals each day.

Variables included in this dataset:

1. **steps**: Number of steps taking in a 5-minute interval (missing values are coded as NA)
2. **date**: The date on which the measurement was taken in YYYY-MM-DD format
3. **interval**: Identifier for the 5-minute interval in which measurement was taken

This dataset is stored in a comma-separated-value (CSV) file and with a total of 17,568 observations.

Next, read this data into a dataframe:

```{r echo=TRUE}
activity<-read.csv("activity.csv", header=TRUE)
totalSteps <- aggregate(
	steps ~ date,
	data = activity,
	sum,
	na.rm = TRUE
)
```

Aggregate relevant data in *totalSteps*.

## What is mean total number of steps taken per day?

Make a histogram of the total number of steps taken daily.

```{r echo=TRUE}
hist(totalSteps$steps)
```

Compute and report the mean and median of total steps taken daily.
```{r echo=TRUE}
meansteps <- mean(totalSteps$steps)
mediansteps <- median(totalSteps$steps)
```

**Mean** total number of steps: `r meansteps`
**Median** total number of steps: `r mediansteps`

## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
